---
title: "EB GAIT"
subtitle: "Evidence Based Gait Analysis Interpretation Tool"
Author: ""
Date: ""

format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    fig-dpi: 600
    embed-resources: true
    theme:
      - cosmo
    grid:
      body-width: 1050px
      sidebar-width: 250px
      margin-width: 0px

html-table-processing: none
    
execute:
    echo: false
    warning: false
    message: false
    error: false
    
params:
  MRN: 481928
  Event_Date: "2024-01-02"
---

```{r}
#| echo: false
#| message: false

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE
)

# BART stuff -----
options(java.parameters = "-Xmx32g")
library(bartMachine, verbose = F)

# Load libraries -----
library(rstanarm, verbose = F)
options(mc.cores = 4)
library(tidyverse, verbose = F)
library(lubridate, verbose = F)
library(patchwork, verbose = F)
library(ggtext, verbose = F)
library(RColorBrewer, verbose = F)
library(glue, verbose = F)
library(gt, verbose = F)
library(gtsummary, verbose = F)
library(ggpmisc, verbose = F)
library(MatchIt, verbose = F)
library(ggh4x, verbose = F)
library(extraDistr, verbose = F)
library(iml, verbose = F)
library(Hmisc, verbose = F)
library(colorspace, verbose = F)
library(MatchIt, verbose = F)

# User Functions -----
margin <- ggplot2::margin
summarize <- dplyr::summarize
select <- dplyr::select
formals(table)$useNA <- "ifany"

# Packages
source("SCRIPTS/read_and_organize.R")
source("SCRIPTS/addnoise.R")
source("SCRIPTS/get_TD_NEW.R")
source("SCRIPTS/build_mod.R")
source("SCRIPTS/get_model_vars.R")
source("SCRIPTS/prop_shap_plot.R")
source("SCRIPTS/prop_detail_plot.R")
source("SCRIPTS/pmatch_plot.R")
source("SCRIPTS/get_outcome_vars.R")
source("SCRIPTS/pmatch_data.R")
source("SCRIPTS/outcome_plots.R")
source("SCRIPTS/p_match_qual.R")
source("SCRIPTS/chkmatchlims.R")
source("SCRIPTS/test_pred_perf.R")
source("SCRIPTS/noisify_side.R")
source("SCRIPTS/getnoisevars.R")
source("SCRIPTS/prop_shap_table.R")
source("SCRIPTS/fmt_shap.R")
source("SCRIPTS/fmt_shapLR.R")
source("SCRIPTS/ptableLR.R")
source("SCRIPTS/getq.R")
source("SCRIPTS/fmt_chk.R")
source("SCRIPTS/get_varlabs.R")
source("SCRIPTS/plot_development.R")

# Make ggplot2 theme -----
theme_mhs <- theme_minimal(base_size = 7) 

theme_mhs <- function(bs){
  theme_minimal(base_size = bs) %+replace%
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = .1, color = "grey85")
    )
}

theme_set(theme_mhs(7))

# For trimming white space around figures -----
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
```

```{r}
#| label: read-td
#| include: false

# Read and organize TD gait data -----
tdavg <- get_TD_NEW()

# Get labels (variable, and pos/neg directions) -----
varlabs <- get_varlabs()

vlabs <- varlabs$Label
names(vlabs) <- varlabs$Variable 
neglabs <- varlabs$negdir
names(neglabs) <- varlabs$Variable
poslabs <- varlabs$posdir
names(poslabs) <- varlabs$Variable
```

```{r}
#| label: "read-organize-data"
#| include: false

# Get clinical data from RRFMP output -----
if(file.exists("DATA/preprocessed_data_2023-12-9.RDS")){
  dat <- readRDS("DATA/preprocessed_data_2023-12-9.RDS")
} else {
  dat0 <- readRDS("DATA/FDpp_2023-11-05.RDS")
  dat <- read_and_organize(dat0)
}

# Get all pre- data (used in GOAL table, etc...) -----
datpre <- readRDS("DATA/FD_2023-11-05.RDS")

```

```{r}
#| label: load-propensity-models

# Note: propensity models are built in EBGAITII.qmd as of 2023-DEC
# List of surgeries -----
surglist <- c("Neural_Rhizotomy", "Rectus_Transfer", "Psoas_Release", "Hams_Lengthening", "Adductor_Release", "Gastroc_Soleus_Lengthening", "Femoral_Derotation_Osteotomy", "Tibial_Derotation_Osteotomy", "DFEO", "Patellar_Advance",  "Foot_and_Ankle_Bone", "Foot_and_Ankle_Soft_Tissue")

# Get probability of surgeries -----
for (s in surglist) {
  load(glue("OUTPUT/Propensity_Models/prop_{s}.RDATA"))
  assign(glue("mod_{s}"), slist[[1]])
  assign(glue("opt_{s}"), slist[[2]])
  mod <- get(glue("mod_{s}"))
}
```

```{r}
#| label: compute-all-propensities
#| cache: true

# Find portion of posterior above threshold -----
ftemp <- function(x, thresh){
  y <- sum(x > thresh) / length(x)
  return(y)
} 

# Get probability of surgeries -----
if(is.null(dat$p_Femoral_Derotation_Osteotomy)){
  for (s in surglist) {
    # cat("Computing Propensity for Surgery ", s, "\n")
    mod <- get(glue("mod_{s}"))
    
    # Get posterior propensity distributions -----
    set.seed(42)
    prop_post <- bart_machine_get_posterior(mod, dat[names(mod$X)])$y_hat_posterior_samples
    dat[[glue("p_{s}")]] <- apply(prop_post, 1, ftemp, thresh = mod$prob_rule_class)
  }
}
```

```{r}
#| label: read-patient-data
#| cache: true

source("SCRIPTS/read_c3d_one.R")
source("SCRIPTS/make_FD_one.R")

# Read in the patient c3d files -----
datc3d <- read_c3d_one(params$MRN)

# Read in patient database info and organize -----
temp <-
  nrow(datpre |> filter(MRN == params$MRN, Event_Date == params$Event_Date))
if(temp == 2){
  FD <- 
    datpre |> 
    filter(MRN == params$MRN)
  
  FDpp  <- 
    dat |> 
    filter(MRN == params$MRN)
  fd <- list("FD" = FD, "FDpp" = FDpp)
} else {
  fd <- make_FD_one(datc3d, params$MRN)
}

# Get visit of interest -----
xpt <- 
  fd$FD %>% 
  filter(
    MRN == params$MRN, 
    year(Event_Date) == year(as.Date(params$Event_Date)),
    month(Event_Date) == month(as.Date(params$Event_Date)),
    day(Event_Date) == day(as.Date(params$Event_Date))
  ) |> 
  read_and_organize()

# Organize, mutate, etc... consistent with historical data -----
ptname <- glue("{xpt$Child_First} {xpt$Child_Last}")[1]
ptdate <- format(as.Date(params$Event_Date), "%b %d, %Y")
```

## General Information

```{r}
#| label: patient-summary

tit <- glue("{ptname} ({params$MRN}), Exam Date: {ptdate}")

tdat <- 
  xpt |> 
  filter(SIDE == "L") |> 
  select(age, dx, dxmod, dxside, GMFCS, FAQ) |> 
  mutate(age = round(age,1)) |> 
  rename_with(
    .fn = ~ vlabs[.],
    .cols = everything()
  )

ptsumm <- 
  tdat |> 
  gt() |> 
  tab_header(tit) |> 
  tab_style(
    style = cell_text(align = "center"),
    locations = list(cells_body(), cells_column_labels())
  ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |> 
  cols_width(
    c(Age, Diagnosis, `Diagnosis Modifier`, GMFCS) ~ px(100),
    `Functional Assessment Questionnaire` ~ 200
  )

ptsumm
```

## Patient History

```{r}
#| label: get-more-patient-info
#| cache: true

source("SCRIPTS/get_patient_info.R")

pi <- get_patient_info(xpt)
```

::: {.panel-tabset .nav-pills}
### Birth 
```{r}
pi$tbirth
```

### Gait
```{r}
pi$tdev
```

### Treatment
```{r}
pi$tsurg
```
:::

## Development

The plots below show the average development of various measurements with age. It is important to keep in mind that many of these exhibit substantial uncertainty (e.g., $\approx\pm 30^\circ$ for Trochanteric Prominence Angle). This uncertainty is not shown.

```{r}
#| label: get-dev-data

devdat <- readxl::read_xlsx("DATA/growth_curves.xlsx")

devdat <- 
  devdat |> 
  rename(vbl = name) |> 
  pivot_longer(-c(vbl, sex), names_to = c("pct", "age"), names_sep = "_") |> 
  mutate(
    age = as.numeric(str_remove(age, "y")),
    name = vlabs[vbl]
  )
  
```


::: {.panel-tabset .nav-pills}
### Height & Weight

```{r}
#| label: plot-ht-wt
#| fig-height: 3.5

# Compute leg length diff
source("SCRIPTS/get_LLD.R")
dat <- get_LLD(dat, "LEG_LENGTH", "LLD")
dat <- get_LLD(dat, "EOS_LegLength", "EOS_LLD")
fd$FD <- get_LLD(fd$FD, "LEG_LENGTH", "LLD")
fd$FD <- get_LLD(fd$FD, "EOS_LegLength", "EOS_LLD")

v <- c("WEIGHT", "HEIGHT", "LLD", "EOS_LLD")
cap = "Weight in Kg, Height/Leg Length in cm"
p <- plot_development(fd, dat, v, v, cap, devdat)
p
```

### Femoral Anteversion & Tibial Torsion

```{r}
#| label: plot-av-tt
#| fig-height: 3.5

v <- c("ANTEVERSION", "EOS_FemTor", "BIMAL", "EOS_Bimal")
cap = "All angles in degrees"
p <- plot_development(fd, dat, v, v, cap, devdat)
p
```

### Motor Control, Mobility, and Gait

```{r}
#| label: plot-gdi-dmc-faqt
#| fig-height: 3.5

v <- c("DMC", "FAQT", "GDI_L", "GDI_R")
vv <- c("DMC", "FAQT", "GDI")
cap = "Dimensionless"
p <- plot_development(fd, dat, v, vv, cap, devdat)
p
```
:::

## Video

```{r}
#| label: get-video
#| include: false

curr.vid <- xpt$Video_File[1]
new.vid <- "DATA/temp.mp4"
new_comp.vid <- "DATA/temp_compressed.mp4"

# Remove existing and copy new -----
unlink(new.vid, force = TRUE)
unlink(new_comp.vid, force = TRUE)

invisible(
  file.copy(curr.vid, new.vid, overwrite = TRUE)
)

# FFmpeg command for compression
ffmpeg_command <-
  paste(
    "ffmpeg -i",
    new.vid,
    "-c:v libx264 -crf 25",
    new_comp.vid
  )

# Execute the FFmpeg command
system(ffmpeg_command)
```

```{r}
#| label: display-compressed-video
#| results: asis

cat("<video width=800 controls>
  <source src=", new_comp.vid, " type=video/mp4>
  </video>")

```

```{r}
#| label: display-original-video
#| results: asis
#| eval: false

cat("<video width=800 controls>
  <source src=", new.vid, " type=video/mp4>
  </video>")

```

## Gait Outcome Assessment List (GOAL) 

Elements of the GOAL survey and physical examination are classified using a subjective mild/moderate/severe scale. Cells are filled according to the the magnitude of deviation ([**none**]{style="color: #e2e2e2"}, [**mild**]{style="color: #E5C9CD"}, [**moderate**]{style="color: #E39AA8"}, [**severe**]{style="color: #D65979"}, [**not applicable**]{style="color: #f7e0a1"}). Missing data are left blank.

::: {.panel-tabset .nav-pills} 

### Gait Outcome Assessment List (GOAL) 

```{r}
#| label: GOAL-survey
source("SCRIPTS/build_goaltbl.R")

# Build goal table -----
t_goal <- build_goaltbl(xpt, datpre)

# Display if GOAL info exists -----
if(!is.null(t_goal)) t_goal

```

### Very Important GOALs

```{r}
#| label: v-important-goals
#| cache: true

source("SCRIPTS/get_goal.R")
source("SCRIPTS/build_importanttbl.R")

# Get goal table data -----
goal <- get_goal(params)

# Build importance table -----
t_imp <- build_importanttbl(goal, params)

# Display if GOAL info exists -----
if(!is.null(t_imp)) t_imp

```
:::


## Physical Examination

Elements of the physical examination are classified using a subjective mild/moderate/severe scale. Cells are filled according to the the magnitude of deviation ([**none**]{style="color: #e2e2e2"}, [**mild**]{style="color: #E5C9CD"}, [**moderate**]{style="color: #E39AA8"}, [**severe**]{style="color: #D65979"}, [**not applicable**]{style="color: #f7e0a1"}). Missing data are left blank. 

```{r}
#| label: physical-exam

source("SCRIPTS/petbl.R")
source("SCRIPTS/orthoseverity.R")
source("SCRIPTS/build_orthotbl.R")
source("SCRIPTS/build_spatbl.R")
source("SCRIPTS/build_strseltbl.R")
source("SCRIPTS/build_foottbl.R")

# Build ortho tables -----
meas <- c(
  "HIP_FLEX", "HIP_EXT", "HIP_ABD_0", "HIP_ABD_90", "POP_ANG_UNI", "POP_ANG_BI",
  "KNEE_EXT", "KNEE_FLEX", "EXTEN_LAG", "PATELLA_ALTA", "ANK_DORS_0", "ANK_DORS_90"
)
tblcon <- build_orthotbl(xpt, meas)

meas <- c(
  "HIP_INT_ROT", "HIP_EXT_ROT", "ANTEVERSION", "EOS_FemTor", 
  "BIMAL", "SEC_TOE", "EOS_Bimal"
)
tbltor <- build_orthotbl(xpt, meas)

# Build neuro tables -----
tblstrsel <- build_strseltbl(xpt)

meas <- c("HIP_FLEX_SPAS", "ADDUCTOR_SPAS","HAMSTRING_SPAS", "RECT_FEM_SPAS",  "PLANTFLEX_SPAS", "POST_TIB_SPAS", "ANKLE_CLONUS")
tblspa <- build_spatbl(xpt, meas)

# Build foot tables -----
meas <- str_subset(names(xpt), "^WB")
tblwb <- build_foottbl(xpt, meas)

meas <- str_subset(names(xpt), "^NWB")
tblnwb <- build_foottbl(xpt, meas)
```

::: {.panel-tabset .nav-pills}
```{r}
#| label: print-PE-tables
#| results: asis
#| eval: true

temp <- c("Torsion", "Contracture", "Motor Control & Strength", "Spasticity",  
          "Weight-Bearing Foot", "Non-Weight-Bearing Foot")
pe <- list(tbltor, tblcon, tblstrsel, tblspa, tblwb, tblnwb)

for (kk in 1:length(temp)) {
  cat("###", temp[kk], "\n\n")
  print(pe[[kk]])
  cat("\n\n")
}
```
:::

## Kinematics

Patient kinematics are shown. Thin lines are all cycles from all trials, heavy lines are the medians. Vertical grid is 15$^\circ$ in all panels. <br>

```{r}
#| label: build-kinematics

source("SCRIPTS/plot_patient_knm.R")
p_kinem_curr <- plot_patient_knm(datc3d, xpt)

source("SCRIPTS/plot_patient_knm_history.R")
p_kinem_hxL <- plot_patient_knm_history(datc3d, fd, xpt, "L")
p_kinem_hxR <- plot_patient_knm_history(datc3d, fd, xpt, "R")

```

::: {.panel-tabset .nav-pills}
```{r}
#| label: print-kinematics
#| results: asis
#| eval: true
#| fig-height: 6.5

plist <- list()
plist[[1]] <- p_kinem_curr
plist[[2]] <- p_kinem_hxL
plist[[3]] <- p_kinem_hxR

temp <- c("Current", "History Left", "History Right")

for (kk in 1:length(temp)) {
  cat("###", temp[kk], "\n\n")
  print(plist[[kk]])
  cat("\n\n")
}
```
:::

```{r}
#| label: add-measurement-error

# Generate likelihood for predictors = measurement with error added -----
zL <- 
  noisify_side(xpt |> filter(SIDE == "L"), 100) |> 
  mutate(SIDE = "L")
zR <- noisify_side(xpt |> filter(SIDE == "R"), 100) |> 
  mutate(SIDE = "R")
zpt <- bind_rows(zL, zR)
```

```{r}
#| label: compute-propensity
#| cache: true

for (s in surglist) {
  set.seed(42)
  mod <- get(glue("mod_{s}"))
  prop_post <- bart_machine_get_posterior(mod, zpt[names(mod$X)])$y_hat_posterior_samples
  zpt[[glue("p_{s}")]] <- apply(prop_post, 1, ftemp, thresh = mod$prob_rule_class)
}
```

## Propensity Profile

The table shows the likelihood of being prescribed surgery based on historical practice standards. The models are exceptionally good at ruling out treatment (negative predictive values \~ 90% or better) and good at ruling in treatment (positive predictive values \~ 60% - 80%) <br>

```{r}
#| label: propensity-prof
#| cache: false

# Build propensity profile -----
propL <- getq(zpt, "L")
propR <- getq(zpt, "R")

# Generate table -----
proptbl <- ptableLR(propL, propR)
proptbl
```

```{r}
#| label: compute-propensity-details
#| include: false
#| cache: true

# Loop over surgeries -----
tall <- list()

# Build detail tables -----
kk <- 0
for (s in surglist) {
  kk <- kk + 1
  mod <- get(glue("mod_{s}"))
  res <- prop_shap_table(sname=s, mod, x=xpt, varlabs)
  tall[[kk]] <- res
}
names(tall) <- surglist
```

```{r}
#| label: format-propensity-details

ptprops <- bind_rows(propL, propR)

# Format detail tables -----
tLR <- list()
kk <- 0
for (s in surglist) {
  kk <- kk + 1
  tLR[[kk]] <- fmt_shapLR(tall, ptprops, s)
}
```

## Propensity Details

The tables show the factors contributing to the historical practice standard. The table is arranged from largest positive contributor ([**indications**]{style="color: #005600"}) to largest negative contributor ([**counter-indications**]{style="color: #841859"}). The top row shows the median (5%, 95%) propensity, color coded from [**high propensity**]{style="color: #005600"} to [**low propensity**]{style="color: #841859"}. The range in propensities accounts for measurement error in the physical examination and three-dimensional gait data. <br> <br>

::: {.panel-tabset .nav-pills}
```{r}
#| label: propensity-details
#| results: asis

for (kk in 1:length(tLR)) {
  cat("###", str_replace_all(surglist[kk], "_", " "), "\n\n")
  print(tLR[[kk]])
  cat("\n\n")
}
```
:::

```{r}
#| label: add-median-propensity-to-patient-data
#| cache: true

# Add median propensity score to patient data -----
medprop <- 
  zpt |> 
  group_by(SIDE) |> 
  summarize(
    across(
      starts_with("p_"),
      ~ median(.))
  )

xpt <- left_join(xpt, medprop)
xL <- xpt |> filter(SIDE == "L")
xR <- xpt |> filter(SIDE == "R")
```

```{r}
#| label: predict-outcome
#| cache: true
#| eval: true

# List of surgeries to predict -----
slist <- 
  xpt %>% 
  select(starts_with("p_")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(name = str_replace_all(name, "p_", "")) %>% 
  pull(name) |> 
  unique() 

# Loop over surgeries -----
p_s <- list()
jj <- 0
for (s in slist) {
  # Get matched data for surgery s -----
  dmatchL <- pmatch_data(s, datx=dat, xside=xL)
  dmatchR <- pmatch_data(s, datx=dat, xside=xR)
  
  
  # Make outcome plot for each outcome variable for surgery s -----
  v <- get_outcome_vars(s)
  poutL <- list()
  poutR <- list()
  kk <- 0
  for (vv in v) {
    kk <- kk + 1
    poutL[[kk]] <-
      outcome_plots(
        match_data = dmatchL,
        s = s,
        outvar = vv,
        xptsd = xL,
        dref = dat,
        vlabs = vlabs,
        neglabs = neglabs,
        poslabs = poslabs
      )
    poutR[[kk]] <-
      outcome_plots(
        match_data = dmatchR,
        s = s,
        outvar = vv,
        xptsd = xR,
        dref = dat,
        vlabs = vlabs,
        neglabs = neglabs,
        poslabs = poslabs
      )
  }
  
  # Assemble outcome plots -----
  pL <- poutL[[1]] / poutL[[2]] / poutL[[3]] / poutL[[4]] +
    plot_annotation(str_replace_all(s, "_", " ")) &
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      plot.margin = margin(c(8, 5, 8, 5))
    )  
  
  pR <- poutR[[1]] / poutR[[2]] / poutR[[3]] / poutR[[4]] +
    plot_annotation(str_replace_all(s, "_", " ")) &
    theme(
      plot.title = element_text(size = 12, face = "bold"),
      plot.margin = margin(c(8, 5, 8, 5))
    )
  
  # Store assembled outcome plots in p_s -----
  p_s[[s]] <- pL | pR
} 
```

## Outcome Predictions

Estimated isolated treatment outcomes for each surgery based on matched historical limbs. [**Treated**]{style="color: #005600"} limbs underwent the highlighted surgery while [**control**]{style="color: #841859"} limbs did not. The mean (point), 50% prediction interval (thick line), and 90% prediction interval (thin line) are shown, along with outcomes of individual matches (tick marks). The effect of the **isolated** treatment on the given outcome measure is described using Cohen's D and conventional thresholds.

Plots show $\Delta$ = Post - Pre [*Changes*]{.underline} for Scale Variables\
Plots show Pre (light) & Post (dark) [*Distributions*]{.underline} for Categorical Variables <br> <br>

::: {.panel-tabset .nav-pills}
```{r}
#| label: plot-outcome-predictions
#| results: asis
#| fig-height: 6.5

for (kk in 1:length(p_s)) {
  cat("###", str_replace_all(names(p_s)[[kk]], "_", " "), "\n\n")
  print(p_s[[kk]])
  cat("\n\n")
}

```
:::
